MSDS6372Project2

### 2021-04-06 MSDS6372Project2.Rmd
update random forest with different trial

#### Objective 2 part 3 - Nonparametric model approach - Random forest
Default RF ntree = 10 and m variables =sqrt(p) seem to work better than ntree=100
```{r}
library(randomForest)
set.seed(1)
# By default use sqrt(p) number variables when building a random forest of classification trees
rf.adult = randomForest(income~.,data=adult)

@@

#Calculate logistic regression complex using cut of 0.425 on test set
cutoff3<-0.425
class.rf <-factor(ifelse(rf.prob>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
confusionMatrix(class.rf,adult.test$income)
# Accuracy : 0.8114 Sensitivity : 0.8118 Specificity : 0.8102

@@

>>>>

#### Objective 2 part 3 - Nonparametric model approach - Random forest
- Default RF ntree = 500 and m variables =sqrt(p) seem to work better than ntree=100
- Try iterate through mtry variable from 1 to 10 using prediction cut off previously discovered - seems mtry = 1 works best on overall accuracy
- After actually fitting random forest with mtry = 1 the AUC = 0.882 (performance is worse compare to default mtry=3). Sensitivity and Specificity also suffers after adjusting the cut off. So Default RF option still wins!
- So Mtry = 3 seems to be the perfect balance for our dataset in order to provide good AUC.
```{r}
library(randomForest)
set.seed(1)
# By default use sqrt(p) number variables when building a random forest of classification trees
rf.adult = randomForest(income~.,data=adult)

@@

#Calculate logistic regression complex using cut of 0.425 on test set
cutoff3<-0.425
class.rf <-factor(ifelse(rf.prob>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
cm = confusionMatrix(class.rf,adult.test$income)
# Accuracy : 0.8114 Sensitivity : 0.8118 Specificity : 0.8102

@@

##### Find best mtry ####
# Use custom loop to find it based on training model
# Using previous discovered cutoff3 = 0.425 
parameter = 10
overall.Accuracy = c()
for(i in 1:parameter){
  set.seed(1)
  rf.adult.temp = randomForest(income~.,data=adult,mtry=i)
  rf.pred.temp = predict(rf.adult.temp,newdata=testdata,type = "prob")
  rf.prob.temp = rf.pred.temp[,2]
  class.rf.temp <-factor(ifelse(rf.prob.temp>cutoff3,">50K.","<=50K."),levels=c("<=50K.",">50K."))
  cm.temp = confusionMatrix(class.rf.temp,adult.test$income)
  overall.Accuracy[i]=cm.temp$overall[1]
}
overall.Accuracy

###### Giving us Mtry = 1 is the best split ####
# Try this option

set.seed(1)
rf.adult2 = randomForest(income~.,data=adult,mtry=1)

# Run randomForest prediction
rf.pred.prob2 = predict(rf.adult2,newdata=testdata,type = "prob")
rf.prob2 = rf.pred.prob2[,2]

# Generate ROC curve to find a cutoff to use
results.rf2 <-prediction(rf.prob2, adult.test$income,label.ordering=c("<=50K.",">50K."))
roc.rf2 = performance(results.rf2, measure = "tpr", x.measure = "fpr")
plot(roc.rf2,colorize = TRUE)
abline(a=0, b= 1)
auc.rf2 <- performance(results.rf2, measure = "auc")
auc.rf2 <- auc.rf2@y.values
text(x = .40, y = .6,paste("AUC = ", round(auc.rf2[[1]],3), sep = ""))
# AUC seem to be the best yet AUC = 0.882
# To balance off performance between high True positive and low False positive we can use around 0.2

#Calculate logistic regression complex using cut of 0.43 on test set
cutoff5<-0.16
class.rf2 <-factor(ifelse(rf.prob2>cutoff5,">50K.","<=50K."),levels=c("<=50K.",">50K."))
confusionMatrix(class.rf2,adult.test$income)
